%%% The main file. It contains definitions of basic parameters and includes all other parts.

%% Settings for single-side (simplex) printing
% Margins: left 40mm, right 25mm, top and bottom 25mm
% (but beware, LaTeX adds 1in implicitly)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
% \openright makes the following text appear on a right-hand page
\let\openright=\clearpage

%% Settings for two-sided (duplex) printing
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{14.2mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\usepackage[utf8]{inputenc}	
\usepackage{tcolorbox}	
\usepackage{float}	
\newtcbox{\inlinecode}{on line, boxrule=0pt, boxsep=0pt, top=2pt, left=2pt, bottom=2pt, right=2pt, colback=gray!15, colframe=white, fontupper={\ttfamily \footnotesize}}	
   	
\usepackage[utf8]{inputenc}	
\usepackage{listings}	
\renewcommand{\lstlistingname}{Figure}% Listing -> Figure	
\renewcommand{\lstlistlistingname}{List of \lstlistingname s}% List of Listings -> List of Figures	
\usepackage{xassoccnt}	
\DeclareCoupledCounters[name=figurelistingsgroup]{figure,lstlisting}	
\makeatletter	
\AtBeginDocument{%	
  \let\c@figure\c@lstlisting	
  \let\thefigure\thelstlisting	
  \let\ftype@lstlisting\ftype@figure % give the floats the same precedence	
}	
\makeatother	
%New colors defined below	
\definecolor{codegreen}{rgb}{0,0.6,0}	
\definecolor{codegray}{rgb}{0.5,0.5,0.5}	
\definecolor{codepurple}{rgb}{0.58,0,0.82}	
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}	
%Code listing style named "mystyle"	
\lstdefinestyle{mystyle}{	
  backgroundcolor=\color{backcolour},   commentstyle=\color{codegreen},	
  keywordstyle=\color{magenta},	
  numberstyle=\tiny\color{codegray},	
  stringstyle=\color{codepurple},	
  commentstyle=\ttfamily,	
  basicstyle=\ttfamily\footnotesize,	
  breakatwhitespace=false,         	
  breaklines=true,                 	
  captionpos=b,                    	
  keepspaces=true,                 	
  numbers=left,                    	
  numbersep=5pt,                  	
  showspaces=false,                	
  showstringspaces=false,	
  showtabs=false,                  	
  tabsize=2	
}	
%"mystyle" code listing set	
\lstset{style=mystyle}	

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Generate PDF/A-2u
\usepackage[a-2u]{pdfx}

%% Character encoding: usually latin2, cp1250 or utf8:
\usepackage[utf8]{inputenc}

%% Prefer Latin Modern fonts
\usepackage{lmodern}

%% Indent the first lines in the section/paragraph	
\usepackage{indentfirst}

%% Further useful packages (included in most LaTeX distributions)
\usepackage{amsmath}        % extensions for typesetting of math
\usepackage{amsfonts}       % math fonts
\usepackage{amsthm}         % theorems, definitions, etc.
\usepackage{bbding}         % various symbols (squares, asterisks, scissors, ...)
\usepackage{bm}             % boldface symbols (\bm)
\usepackage{graphicx}       % embedding of pictures
\usepackage{fancyvrb}       % improved verbatim environment
%\usepackage{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
\usepackage[nottoc]{tocbibind} % makes sure that bibliography and the lists
			    % of figures/tables are included in the table
			    % of contents
\usepackage{dcolumn}        % improved alignment of table columns
\usepackage{booktabs}       % improved horizontal lines in tables
\usepackage{paralist}       % improved enumerate and itemize
\usepackage{xcolor}         % typesetting in color
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{adjustbox}	
\usepackage{wrapfig}	
\usepackage[htt]{hyphenat}	
\usepackage[T1]{fontenc}	
\usepackage[final]{microtype}	
\usepackage{listings} %For code in appendix	
\lstset	
{ %Formatting for code in appendix	
    language=Python,	
    basicstyle=\footnotesize,	
    numbers=left,	
    stepnumber=1,	
    showstringspaces=false,	
    tabsize=2,	
    breaklines=true,	
    breakatwhitespace=false,	
    keywordstyle=\ttfamily,	
    stringstyle=\ttfamily,	
    identifierstyle=\ttfamily	
}

\usepackage[style=numeric,backend=biber,sorting=none]{biblatex}
\setcounter{biburllcpenalty}{7000}
\setcounter{biburlucpenalty}{8000}
\addbibresource{bibliography.bib}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Basic information on the thesis

% Thesis title in English (exactly as in the formal assignment)
\def\ThesisTitle{Data Lineage Analysis Service for Embedded Code}

% Author of the thesis
\def\ThesisAuthor{Michal Jurčo}

% Year when the thesis is submitted
\def\YearSubmitted{2023}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% or a full name of a department outside MFF)
\def\Department{Department of Distributed and Dependable Systems}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

% Thesis supervisor: name, surname and titles
\def\Supervisor{doc. RNDr. Pavel Parízek, Ph.D.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Department of Distributed and Dependable Systems}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Software and Data Engineering}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, a person who lent the software, etc.)
\def\Dedication{%
This way I would like to thank my supervisor,  doc. RNDr. Pavel Parízek, for his guidance, help and motivation while working on this thesis.\\
Thanks should also go to RNDr. Lukáš Hermann for his invaluable assistance regarding
MANTA Flow platform and mentorship.\\
I am especially grateful to my fiancée, Kristýna, who has shown her never-ending support and love throughout my studies that helped me get this far.\\
Last but not least, I want to thank my family and my cats for always providing comfort when I needed it.
}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
Data integration tools often use embedded code for data manipulation tasks. Popular examples of such tools include AWS Glue data integration service, Databricks platform, Snowflake data cloud or SQL Server Integration Services (SSIS). Embedded code is typically written in programming languages such as Python, Java, C\# or JavaScript. Manta Flow is an automated platform that can analyze data lineage in database models, data pipelines of data integration tools, and in application source code, but it lacks the ability to analyze embedded code.\\
In this work, we analyzed potential ways to extend the capabilities of Manta Flow with the ability to analyze data lineage in embedded code. We created a general design of a reusable Embedded Code Service that leverages the existing potential of data flow analysis of source code, and uses it to analyze embedded code. We implemented a specialization of this service for the Python programming language, and to demonstrate its usefulness, we designed and implemented a prototype of data lineage scanner for AWS Glue data integration service. This scanner extensively uses the service to analyze data lineage in embedded Python scripts.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces
\def\Keywords{%
{data lineage} {data flow} {embedded code} {python}
}

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
\hypersetup{unicode}
\hypersetup{breaklinks=true}
\hypersetup{hidelinks}

% Definitions of macros (see description inside)
\include{macros}

% Title page and various mandatory informational pages
\begin{document}
\include{title}

%%% A page with automatically generated table of contents of the master thesis

\tableofcontents

%%% Each chapter is kept in a separate file
\include{preface}
\include{chap01}
\include{chap02}
\include{chap03}
\include{chap04}
\include{chap05}
\include{epilog}

%%% Bibliography
%\include{bibliography}
\printbibliography

%%% Figures used in the thesis (consider if this is needed)
\listoffigures

%%% Tables used in the thesis (consider if this is needed)
%%% In mathematical theses, it could be better to move the list of tables to the beginning of the thesis.
%\listoftables

%%% Abbreviations used in the thesis, if any, including their explanation
%%% In mathematical theses, it could be better to move the list of abbreviations to the beginning of the thesis.
%\chapwithtoc{List of Abbreviations}

%%% Attachments to the master thesis, if any. Each attachment must be
%%% referred to at least once from the text of the thesis. Attachments
%%% are numbered.
%%%
%%% The printed version should preferably contain attachments, which can be
%%% read (additional tables and charts, supplementary text, examples of
%%% program output, etc.). The electronic version is more suited for attachments
%%% which will likely be used in an electronic form rather than read (program
%%% source code, data files, interactive charts, etc.). Electronic attachments
%%% should be uploaded to SIS and optionally also included in the thesis on a~CD/DVD.
%%% Allowed file formats are specified in provision of the rector no. 72/2017.
\appendix
\chapter{Attachments}

\section{User Documentation}

To run metadata extraction and data flow analysis of AWS Glue scanner, you are required to have a computer with:
\begin{itemize}
    \item runtime environment for Java 11 or newer installed,
    \item Manta Flow, version R40 or newer installed and configured.
\end{itemize}
Manta Flow requires a license to be successfully configured. Licenses are available only to Manta customers and developers, which might be a problem for the reader.
\par
The installer for Manta Flow contains a wizard for its configuration. After the installation is successfully finished, Manta Flow can be launched.
\par
After logging into Manta Flow Admin UI, it is necessary to create AWS Glue connection, which is listed under Data Integration tools. The connection needs to be named uniquely and requires AWS Glue credentials. The user needs to generate programmatic credentials for the AWS user credentials that will be used to access AWS Glue. It is also possible to specify a regular expression which defines which ETL jobs shall be analyzed.
\par
With connection set up, we need to create a workflow for extraction analysis. In process manager, create a new workflow and drag Extraction and Analysis into the workflow. They should contain your AWS Glue connection. When the workflow is created, execute it and watch the process manager to see when it completes.
\par
After the workflow finished, you may log into Manta Flow Viewer to visualize the analyzed data lineage. In the Viewer, select the latest revision and choose, which objects you would like to visualize. Then, click \textit{Visualize} button and explore the data lineage.

\section{Contents of the Attachment}

The list below describes the contents of the attachment:
\begin{itemize}
    \item The \textit{source-code} folder contains the source code developed within this work. There are only classes and other files produced in their entirety as a part of this work. Note that the source code of related Manta Flow components is not included. The source code is provided in the hierarchical structure as it appears in the development source code management tool.
    \item The \textit{diff} folder contains two files in \textit{diff} format (format used to store differences between two files) that shows the changes that were made to files not included in the \textit{source-code} folder, but were a part of this work. The changes include the refactoring of Python scanner components for the purposes of integration with Embedded Code Service.
    \item The \textit{tex-source} contains the source files used to generate PDF version of this text. It contains the original version of the diagrams and pictures shown in figures in this work. Should any of them be difficult to read, you can examine them in better resolution.
\end{itemize}

\openright
\end{document}
